---
title: "Homework 0: Pre-Course Assignment"
author: "Megan Marziali"
date: "January 3, 2021"
output: github_document
---

This document contains a solution for the pre-course assignment for EPID P8451.

```{r setup, message=FALSE}
library(tidyverse)
library(arsenal)
```

## Question 1: Summary Table

The code chunk below imports and cleans the data.

```{r, message=FALSE}
bc_data = 
  read_csv("./data/bcdata_Assignment1.csv",
           na = "") %>% 
  janitor::clean_names()
```

The dataset contains 10 variables and 116 observations.

The code chunk below creates a summary table of the quantitative features of the dataset.

```{r, results="asis"}
table_one = tableby(~ age + bmi + glucose + insulin
                    + homa + leptin + adiponectin 
                    + resistin + mcp_1, data = bc_data,
                    numeric.stats = c("mean","median", "range"))

summary(table_one)
```

## Question 1: Recoding BMI

The code chunk below will recode BMI into the WHO defined categories.

```{r}
bc_data %>% 
  mutate(
    bmi_cat = case_when(bmi < 16.5 ~ "severely underweight", 
                        bmi < 18.5 ~ "underweight",
                        bmi >= 18.5 & bmi < 25 ~ "normal",
                        bmi >= 25 & bmi < 30 ~ "overweight",
                        bmi >= 30 & bmi < 35 ~ "obese class I",
                        bmi >= 35 & bmi < 40 ~ "obese class II",
                        bmi >= 40 ~ "obese class III"))
```


